### 2024-10-16 plroba_latex.Rmd
weiter

---
title: "Insert Your Title Here"
shorttitle: "Short Title for Headers"
author:
  - Ime1 Prezime1 \affil{1}\comma\corrauth
shortauthor: "Prezime1, Prezime2 & Prezime3"
address: |
  \affilnum{1}\ Sveučilište..., Fakultet ili Odjel..., Adresa1, Zemlja1 \href{mailto:abc@abc.com}{\faEnvelopeO}
  \newline
  \affilnum{2}\ Sveučilište..., Fakultet ili Odjel..., Adresa2, Zemlja2
date: "`r format(Sys.Date(), '%d %B %Y')`"
thisnumber: "x"
thisyear: "202x"
thisvolume: "x"
datereceived: "xxxx xx, 202x"
dateaccepted: "xxxx xx, 202x"
dateavailable: "xxxx xx, 202x"
doinumber: "10.62366/crebss.202x.x.00x"
type: "XXX XXX"
JEL: "XXX, XXX"
keywords: "izraz1, izraz2,..."
abstract: |
  Sažetak ne smije biti dulji od 200 riječi, ali ne manji od 150. Sažetak treba ilustrirati sadržaj članka s jasnim ciljevima, metodologijom, rezultatima i doprinosom. Članak se može klasificirati kao (i) izvorni znanstveni rad, (ii) pregledni rad ili (iii) prethodno priopćenje, ovisno o prijedlozima recenzenata.
output:
  pdf_document:
    template: template.tex
    keep_tex: true
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.