### 2024-12-23 plroba_latex.Rmd
weiter

---
title: "Test Document"
shorttitle: "Test"
author:
address: |
  \affilnum{1}\ Sveučilište..., Fakultet ili Odjel..., Adresa1, Zemlja1 \href{mailto:abc@abc.com}{\faEnvelopeO}
keywords: "test, latex, RMarkdown"
abstract: |
  This is a test abstract to verify LaTeX compilation with the custom class.
output:
  pdf_document:
    template: template.tex
    keep_tex: true
    latex_engine: pdflatex
---

# Introduction

This is a test document. Here's a citation: \citep{shoukralla:23}.

\begin{equation} \label{eq:test}
E = mc^2
\end{equation}

Referencing the equation: \eqref{eq:test}.

# Conclusion

Test conclusion.

# References

\begin{thebibliography}{99}

\bibitem[Shoukralla and Ahmed(2023)]{shoukralla:23}
Shoukralla, E. S. and Ahmed, B. M. (2023). Barycentric Lagrange Interpolation Methods for Evaluating Singular Integrals. *Alexandria Engineering Journal*, 69, 243–253. \href{https://doi.org/10.1016/j.aej.2022.12.005}{doi:10.1016/j.aej.2022.12.005}

\end{thebibliography}