### 2024-12-23 plroba_latex.tex

\documentclass{crebsshr}

%%%%%%%%% Journal Info -- DO NOT CHANGE %%%%%%%%%
\setcounter{page}{1}
\setcounter{secnumdepth}{4}
\renewcommand\thisnumber{x}
\renewcommand\thisyear {202x}
\renewcommand\thisvolume{x}
\renewcommand\datereceived{xxxx xx, 202x}
\renewcommand\dateaccepted{xxxx xx, 202x}
\renewcommand\dateavailable{xxxx xx, 202x}
\renewcommand\doinumber{10.62366/crebss.202x.x.00x}
\renewcommand\type{XXX XXX}
\renewcommand\JEL{XXX, XXX}
%%%%%%%%% End Journal Info   %%%%%%%%%

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{fontawesome}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{float}  % For [H] placement
\usepackage{booktabs}  % For professional tables
\usepackage{subcaption}  % For subfigures
\usepackage{geometry}  % Adjust margins if needed

% Custom commands from the blueprint
\newcommand{\affilnum}[1]{\textsuperscript{#1}}

% Begin document
\begin{document}

% The title, author, abstract, etc., will be inserted by R Markdown

\section{Introduction}\label{introduction}

This is a test document. Here's a citation: \citep{shoukralla:23}.

\begin{equation} \label{eq:test}
E = mc^2
\end{equation}

Referencing the equation: \eqref{eq:test}.

\section{Conclusion}\label{conclusion}

Test conclusion.

\section{References}\label{references}

\begin{thebibliography}{99}

\bibitem[Shoukralla and Ahmed(2023)]{shoukralla:23}
Shoukralla, E. S. and Ahmed, B. M. (2023). Barycentric Lagrange Interpolation Methods for Evaluating Singular Integrals. *Alexandria Engineering Journal*, 69, 243–253. \href{https://doi.org/10.1016/j.aej.2022.12.005}{doi:10.1016/j.aej.2022.12.005}

\end{thebibliography}

% Bibliography (if using BibTeX)
\bibliographystyle{plain}
\bibliography{references}

\end{document}